---
title: "Task 2: Political Affiliation and Environmentalist Identification"
author: "Katheryn Moya"
date: "11/24/2021"
output: html_document
---

The data used in this visualization was collected and made available by Chelsea Batavia and colleagues at the Andrews Forest Long Term Ecological Research Program. The data reflects a survey of 1,331 respondents across the United States in August 2017 surrounding the influences on charitable given for conservation. The survey data is of two types - information about respondents and their attitude towards seven manipulated textual messages. Comparisons between political orientation (liberal or conservative) and whether or not a person considers themselves an environmentalist occurs through a chi square analysis.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Attach necessary packages 

library(here)
library(tidyverse)
library(broom)
library(janitor)
library(kableExtra)
```



```{r, include = FALSE}

# Read in data

cons_survey <- read_csv(here("data","conservation_survey.csv"))

# Simplify dataframe for ENV_ID and POL_OR

cons_simple <- cons_survey %>% 
  select("ENV_ID", "POL_OR") %>% 
# Add a new column for political party and exclude all other political affiliations
  filter(POL_OR %in% c("1","2","3","5","6","7")) %>%
  mutate(political_party = case_when(
    POL_OR == "1" ~ "Conservative",
    POL_OR == "2" ~ "Conservative",
    POL_OR == "3" ~ "Conservative",
    POL_OR == "5" ~ "Liberal",
    POL_OR == "6" ~ "Liberal",
    POL_OR == "7" ~ "Liberal")) %>% 
  # Change observation names for ENV_ID for clearer table
  mutate(ENV_ID = case_when(
    ENV_ID == "1" ~ "Yes",
    ENV_ID == "2" ~ "No",
    ENV_ID == "3" ~ "Unsure"))
```



```{r, include = FALSE}
# Find counts and proportions of “liberal” and “conservative” survey respondents who responded “Yes”, “No” and “Unsure” to the statement “I consider myself an environmentalist.” 

cons_simple %>% 
  count(political_party, ENV_ID)

cons_counts <- cons_simple %>% 
  tabyl(political_party, ENV_ID)

# Calculate proportions
cons_proporations <- cons_counts %>% 
  adorn_percentages() %>%  
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()
```



```{r}
# Create a finalized table

cons_proporations %>%
  kbl(caption = "**Table 1.** Political Affilation and response to the statement 'I consider myself an environmentalist'",
      col.names = c("",
                    "No",
                    "Yes",
                    "Unsure")) %>%
 kable_classic(full_width = FALSE, html_font = "Cambria") %>% 
  row_spec(0, align = "center", bold = TRUE, color = "#191970", background = "#eee5de") %>% 
   row_spec(1, align = "center", bold = FALSE, color = "white", background = "#f08080") %>% 
    row_spec(2, align = "center", bold = FALSE, color = "#191970", background = "#bfefff")
```



```{r, include = FALSE}

# Convert counts table to a contingency table
cons_ct <- cons_counts %>% 
  column_to_rownames(var = "political_party")
```



```{r, include = FALSE}
# Perform chi-square test
cons_x2 <- chisq.test(cons_ct)
cons_x2

cons_tidy <- tidy(cons_x2)
```
There is a significant association between political affiliation and if a person considers themselves an environmentalist ($\chi$^2^(`r cons_tidy$parameter`) = `r round(cons_tidy$statistic, 2)`, *p* = `r ifelse(cons_tidy$p.value < 0.001, paste("< 0.001"), paste("=", cons_tidy$p.value))`). There is sufficient evidence to retain the alternative hypothesis that political affiliation and environmentalist identification are not independent of each other.


**Data citation:** Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e 
