---
title: "Task 3: Female Lizard Lengths"
author: "Katheryn Moya"
date: "11/27/2021"
output: html_document
---

The data used in this visualization was collected and made available by David Lightfoot and colleagues at the Jornada Basin Long Term Ecological Research Program in southern New Mexico. The data contains  includes 14 variables for 4087 individual observations. Observations represent trapped lizards and the date, location, species, sex, age, total length, and tail condition. Following the exploratory data visualization, length comparisons between female lizards across species (Desert Grassland lizard (*Cnemidophorus uniparens*) and Colorado Checkered lizard (*Cnemidophorus tessalatus*)) through two-sample t-tests using a significance level ($\alpha$) of 0.05 throughout. Differences between groups are described by Cohen's *d* effect size.

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Attach packages

library(here)
library(tidyverse)
library(janitor)
library(kableExtra)
library(effsize)
```


```{r}
# Read in data

lizards <- read_csv(here("data","lizards.csv"))
```

```{r}
# Create simplified data frame

female_liz <- lizards %>% 
  select("spp", "sex", "total_length") %>% 
#filter for female and CNUN and CNTE
  filter(sex == "F") %>% 
  filter(spp %in% c("CNUN","CNTE")) %>% 
# create new column with full name 
  mutate(spp_full = case_when(
    spp == "CNUN" ~ "Desert Grassland",
    spp == "CNTE" ~ "Colorado Checkered"))

```



```{r, fig.cap = "**Figure 1.** Female lizard total lengths (in mm) across species (Desert Grassland lizard (*Cnemidophorus uniparens*) and Colorado Checkered lizard (*Cnemidophorus tessalatus*)) at the Jornada Basin Long Term Ecological Research Program."}
# Data visualization
# compare lengths of female lizards for the two species

ggplot(data = female_liz, aes(x = spp_full, y = total_length)) +
  geom_boxplot(aes(fill = spp)) +
  labs(title = "Female Lizard Lengths in Jornada Basin LTER, New Mexico",
       y = "Total length (mm)",
       x = "Species",
       fill = "Species")+
  scale_x_discrete(labels=c("Colorado Checkered lizard","Desert Grassland lizard")) +
  scale_fill_manual(values = c("coral3", "darkseagreen3"))+
  scale_color_manual(values = c("coral2", "darkseagreen2")) +
  theme_bw()
```


```{r}

# Finalized summary table
# Create data set with stats

female_liz_stats <- female_liz %>% 
  group_by(spp_full) %>% 
  summarize(mean_total_length = round(mean(total_length, na.rm = TRUE), digits = 2),
            sd_total_length = round(sd(total_length, na.rm = TRUE), digits = 2),
            sample_size = n())

# Create a finalized summary table

female_liz_stats %>%
  kbl(caption = "**Table 1.** Total Length Distribution between female Desert Grassland and Colorado Checkered lizards in Jornada Basin LTER",
      col.names = c("Species",
                    "Mean Total Length (mm)",
                    "Standard Deviation (mm)",
                    "Sample Size (n)")) %>%
 kable_classic(full_width = TRUE, html_font = "Cambria") %>% 
   row_spec(0, align = "center", bold = TRUE, color = "#8b5f65", background = "	#ffec8b") %>% 
  row_spec(1:2, align = "center", bold = FALSE, color = "#8b5f65", background = "#fafad2")
```
There is a significant difference in total length difference in mean total length between female lizards for the two species. The actual mean difference in total length between these two species of female lizards is 97.29 mm, with a % difference in total length of 23.5%. Through a Cohen’s d analysis, there is a large effect size, signifying that there is an important difference between the total lengths of female lizards across species.


```{r, include = FALSE}
# make exploratory graphs to explore assumptions of hypothesis test

ggplot(data = female_liz, aes(x = total_length)) +
  geom_histogram(bins = 15) +
  facet_wrap( ~ spp) +
  aes(fill = spp) 

ggplot(data = female_liz, aes(sample = total_length)) +
  geom_qq() +
  facet_wrap( ~ spp) +
  aes(color = spp)
```

```{r, include= FALSE}
# create vectors for CNUN and CNTE length for two-sample t-test

cnun_length <- female_liz %>% 
  filter(spp == "CNUN") %>% 
  drop_na(total_length) %>% # removing weights with NA value
  pull(total_length)

cnte_length <- female_liz %>% 
  filter(spp == "CNTE") %>% 
  drop_na(total_length) %>% 
  pull(total_length)

# two-sample t-test

female_liz_t <- t.test(cnun_length, cnte_length)

female_liz_t

# Cohen's d
female_liz_cohen <- cohen.d(cnun_length, cnte_length)

female_liz_cohen
```
**Data citation:** Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e 


